Resource Adequacy and Operational Security Interaction in the EPOC 2030-50 Project

### 2022-06-14 papers/main.tex
Add more description of data

@@

\section{Description of a Deterministic Unit Commitment Model with Probabilistic Reserve (DUC-PR) constraints}

@@

\begin{itemize}
    \item $G$ - Generators
    \item $G_n$ - Generators located at node $n$
    \item $GD$ - Dispatchable / thermal / conventional generators
    \item $GR$ - Renewable / variable generators
    \item $N$ - Nodes or buses in the network
    \item $B$ - Lines or branches
    \item $T$ - Time steps / intervals / slices
    \item $L^+$ - Upward reserve levels
    \item $L^-$ - Downward reserve levels
\end{itemize}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/imbalance_range}
    \caption{Range of possible imbalances for all hours of day 309.\label{fig:imbalance_range}}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/convex_hull_TIHANGE 1_TIHANGE 2}
    \caption{Illustration of convex hull constraints on $d_{nlt}^{L^+}$ and $d_{nlt}^{L^-}$. Blue circles are imbalance scenarios, shaded area is the convex hull, $d_{nt}^{min}$ and $d_{nt}^{max}$ are the x limits for TIHANGE 1 and the y limits for TIHANGE 2. \label{fig:convex_hull_tihange}}
\end{figure}

@@

\section{Input data - stylised Belgian grid with a large amount of variable renewable energy sources}

\todo[inline]{For now, see README.md}

\section{Analysis of DUC-PR model results}

This analysis relates only to the 309th day of the year (November 5th).

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base/dispatch}
        \caption{Economic dispatch model}
    \end{subfigure}
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true/dispatch}
        \caption{Unit commitment model}
    \end{subfigure}
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true/dispatch}
        \caption{Unit commitment model with day ahead network constraints}
    \end{subfigure}
    \caption{Dispatch schedules with and without operating reserves.\label{fig:dispatch_base}}
\end{figure}

\subsection{Probabilistic operating reserves with no reserve activation network constraints}

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_RSV=0.0/dispatch}
        \caption{No UC, no DANet, no PSCD, RSL is 0.0}
    \end{subfigure}
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_RSV=0.0/dispatch}
        \caption{UC, no DANet, no PSCD, RSL is 0.0}
    \end{subfigure}
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true_RSV=0.0/dispatch}
        \caption{UC, DANet, no PSCD, RSL is 0.0}
    \end{subfigure}
    \caption{Dispatch schedules with and without operating reserves.\label{fig:dispatch_simple_reserves}}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_timeseries_idx=56}
        \caption{}
    \end{subfigure}
    \newline
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_abs_idx=56}
        \caption{}
    \end{subfigure}
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=56}
        \caption{}
    \end{subfigure}
    \caption{Load shedding distribution in time (top figure, aggregated over all nodes) and space (aggregated over the entire day) for UC, DANet, LM = 1.5 and $RSL = 0.5$.}
\end{figure}

>>>>

\author{Sebastian Gonzato}

@@

The rest of this document is organised as follows. \secref{sec:model} describes the \ac{DUCPR} model used to simulate power system operations and obtain day ahead adequacy and very approximate real time operational security indicators, namely scheduled load shedding and reserve shedding. \secref{sec:data} describes the stylised Belgian system analysed. \secref{sec:initial_results} analyses a single day in order to identify sources of adequacy and security issues, while \secref{sec:adequacy_vs_security} analyses the trade-off between these two.

\section{Description of a Deterministic Unit Commitment Model with Probabilistic Reserve (DUC-PR) constraints} \label{sec:model}

@@

\begin{itemize}
    \item $G$ - Generators
    \item $G_n$ - Generators located at node $n$
    \item $GD$ - Dispatchable / thermal / conventional generators
    \item $GR$ - Renewable / variable generators (\ac{VRES}).
    \item $N$ - Nodes or buses in the network
    \item $B$ - Lines or branches
    \item $T$ - Time steps / intervals / slices
    \item $L^+$ - Upward reserve levels
    \item $L^-$ - Downward reserve levels
\end{itemize}

@@

\begin{figure}[ht]
    \centering
    % \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/imbalance_range}
    \caption{Range of possible imbalances for all hours of day 309.\label{fig:imbalance_range}}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    % \includegraphics[width=0.9\textwidth]{../plots/nodal_imbalance_investigation/convex_hull_TIHANGE 1_TIHANGE 2}
    \caption{Illustration of convex hull constraints on $d_{nlt}^{L^+}$ and $d_{nlt}^{L^-}$. Blue circles are imbalance scenarios, shaded area is the convex hull, $d_{nt}^{min}$ and $d_{nt}^{max}$ are the x limits for TIHANGE 1 and the y limits for TIHANGE 2. \label{fig:convex_hull_tihange}}
\end{figure}

@@

\section{Input data - stylised Belgian grid with a large amount of variable renewable energy sources} \label{sec:data}

All data used for this exercise can be found at: \href{https://www.dropbox.com/sh/mdvmc082gwng0tr/AABRyc3fZpxAFycmUfZmh8Csa?dl=0}.

The grid and resource data used for this interaction is inspired by the case study in \cite{Belderbos2020}, omitting the gas network and power to gas technologies. This data gives a stylised Belgian system with a high penetration (~8s0\%) of \ac{VRES}, which includes 75.7 GW of solar PV, 7.0 GW of Onshore Wind and 7.3 GW of Offshore Wind. The grid consists of 46 nodes connected by 69 lines. The total amount of conventional thermal generation is 9.9 GW, with all generators based on new \ac{CCGT} units apart from one new \ac{OCGT} unit at Lixhe. The nuclear generators of Doel and Tihange are therefore replaced with new \ac{CCGT} units.

The residual load time series, aggregated across all nodes, for this system is plotted in \todo[inline]{Plot this}.

A complete overview of the data is given in \cite{Belderbos2020}, Appendix B.

\section{Analysis of \acl{DUCPR} model results} \label{sec:initial_results}

This analysis relates only to the 309th day of the year (November 5th). The aim is to investigate what are the causes of load shedding.

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base/dispatch}
        \caption{Economic dispatch model}
    \end{subfigure}
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true/dispatch}
        \caption{Unit commitment model}
    \end{subfigure}
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true/dispatch}
        \caption{Unit commitment model with day ahead network constraints}
    \end{subfigure}
    \caption{Dispatch schedules with and without operating reserves.\label{fig:dispatch_base}}
\end{figure}

\subsection{Probabilistic operating reserves without reserve activation network constraints}

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_RSV=0.0/dispatch}
        \caption{No UC, no DANet, no PSCD, RSL is 0.0}
    \end{subfigure}
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_RSV=0.0/dispatch}
        \caption{UC, no DANet, no PSCD, RSL is 0.0}
    \end{subfigure}
    \begin{subfigure}[t]{0.7\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../data/sims/main_model_runs/309/base_UC=true_DANet=true_RSV=0.0/dispatch}
        \caption{UC, DANet, no PSCD, RSL is 0.0}
    \end{subfigure}
    \caption{Dispatch schedules with and without operating reserves.\label{fig:dispatch_simple_reserves}}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_timeseries_idx=56}
        \caption{}
    \end{subfigure}
    \newline
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_abs_idx=56}
        \caption{}
    \end{subfigure}
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        % \includegraphics[width=\textwidth]{../plots/main_model_runs/ls_per_node_idx=56}
        \caption{}
    \end{subfigure}
    \caption{Load shedding distribution in time (top figure, aggregated over all nodes) and space (aggregated over the entire day) for UC, DANet, LM = 1.5 and $RSL = 0.5$.}
\end{figure}

\section{Day ahead adequacy and real time operational security trade offs} \label{sec:adequacy_vs_security}
